---
title: "Titanic Dataset Summary"
format: html
editor: visual
---

## Introduction

This report summarizes the Titanic dataset, offering insights into passenger demographics and survival rates from the tragic sinking of the RMS Titanic in 1912.

### Data Loading and Preparation

```{r}
library(ggplot2) ; library(dplyr)

# Load the Titanic dataset
dt <- read.csv("titanic.csv")

# Convert relevant columns to factors
dt[,c(2:5,7:9,11,12)] <- lapply(dt[,c(2:5,7:9,11,12)], as.factor)
head(dt)
```

### Summary of the Data Set

```{r}
# Display structure and summary of the dataset
str(dt)
summary(dt)
```

### Handling Missing Values

```{r}
# Remove rows with NA values
dt <- na.omit(dt)
summary(dt)
```

### Survival Rates:

```{r}
# Calculate overall survival rates
survival_rate <- dt %>% group_by(Survived) %>% summarize(Count = n())

# Plot overall survival rates
ggplot(survival_rate, aes(x = Survived, y = Count)) +
  geom_bar(stat = "identity") +
  labs(x = "Survived", y = "Count", title = "Survival Rates on the Titanic")
```

### Survival Rate by Gender

```{r}
# Calculate survival rates by gender
gender_survival <- dt %>% group_by(Sex, Survived) %>% summarize(Count = n()) %>%
  mutate(Survival_Rate = Count / sum(Count))

# Plot survival rates by gender
ggplot(gender_survival, aes(x = Sex, y = Survival_Rate, fill = Survived)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Gender", y = "Survival Rate", fill = "Survived", title = "Survival Rate by Gender") +
  scale_y_continuous(labels = scales::percent)
```

### Survival Rate by Passenger Class

```{r}
# Calculate survival rates by passenger class
class_survival <- dt %>% group_by(Pclass, Survived) %>% summarize(Count = n()) %>%
  mutate(Survival_Rate = Count / sum(Count))

# Plot survival rates by passenger class
ggplot(class_survival, aes(x = Pclass, y = Survival_Rate, fill = Survived)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Passenger Class", y = "Survival Rate", fill = "Survived", title = "Survival Rate by Passenger Class") +
  scale_y_continuous(labels = scales::percent)
```

### Survival Rate by Age Groups

```{r}
# Create age groups
dt <- dt %>%
  mutate(Age_Group = case_when(
    Age < 18 ~ "<18",
    Age >= 18 & Age < 30 ~ "18-29",
    Age >= 30 & Age < 40 ~ "30-39",
    Age >= 40 & Age < 50 ~ "40-49",
    Age >= 50 & Age < 60 ~ "50-59",
    Age >= 60 ~ "60+",
    TRUE ~ "Unknown"
  ))

# Calculate survival rates by age group
age_survival <- dt %>% group_by(Age_Group, Survived) %>%
  summarize(Count = n()) %>% mutate(Survival_Rate = Count / sum(Count))

# Plot survival rates by age group
ggplot(age_survival, aes(x = Age_Group, y = Survival_Rate, fill = Survived)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Age Group", y = "Survival Rate", fill = "Survived", title = "Survival Rate by Age Group") +
  scale_y_continuous(labels = scales::percent)
```

## Conclusion

This report analyzed the Titanic dataset, revealing significant insights into survival rates based on gender, passenger class, and age groups.
